Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-01-27 Grassmannian_cells.tex
Section 5
I need to clean up the notation and to say something concerning
compatibility in the main proof.

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma \ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section \ref{Lifting}.
The next step is to associate a $2$-quiver with certain lifts of (truncated) preprojective representation $\tilde P_m^{\uj}$ for $\uj\in I_{n-1}$ whose strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians. Recall that the cells of the quiver Grassmannians attached to $P_m^{\uj}:= F_{K(n)}\tilde P_m^{\uj}$ are in one-to-one correspondence with those attached to the lift $\tilde P_m^{\uj}$. Moreover, the quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ attached to truncated preprojective representations only depend on $\tbfe$ and $\udim P_m^V$. In particular, in view of Theorem \ref{thm:truncpp}, these $2$-quivers can be seen as being attached to any (truncated) preprojective of this dimension.

Thus it suffices to consider the lifted truncated preprojectives $\tilde P_m^{[k]}:=\tilde P_m^{(1,\ldots,k)}$ for $m\geq 2$ and $0\leq k\leq n-1$. We define $P_m^{[k]}:=F_{K(n)}\tilde P_m^{[k]}$.

@@

$$ \Hom(P_{m-1},\tau P_m^{[k]})\cong\bigoplus_{i=k+1}^n\Hom(\alpha_{i}^{(-1)^{m}}\tilde P_{m-1},\tau\tilde P_{m}^{[k]})\cong\bigoplus_{i=k+1}^n\Ext(\tilde P_m^{[k]},\alpha_{i}^{(-1)^{m}}\tilde P_{m-1})\cong\Ext(P_m^{[k]},P_{m-1})$$.

@@

 This basis recursively gives rise to the following $2$-quiver of $P_3^{[k]}$ for $k=n-1,\ldots,0$ where we drop the indices of the vertices in abuse of notation. Note that the middle vertex of $\tilde P_3^{[n-1]}$ is $(2,\alpha_n^{-1})$ and that all $2$-arrows end in the vertex corresponding to $(1,1)$ of $\widetilde{K(n)}$.

The representation $\tau\tilde P_3^{[k]}$ is given by 
\[\xymatrix@R10pt@C20pt{&&(2,\alpha_{k+1}^{-1})\ar_{\alpha_{k+1}}[lld]\\ (1,1) &&\vdots\\&&(2,\alpha_n^{-1})\ar_{\alpha_{n}}[llu]}.\]
Note that the image of the homomorphism of the corresponding exact sequence induced by the respective basis elements is the subquiver $(2,\alpha_i^{-1})\xrightarrow{\alpha_i} (1,1)$ for $i=k+1,\ldots, n$. Thus in this case the $2$-arrow connects the $2$-quiver of $\tilde P_3^{[k]}$ to the image of the corresponding homomorphism $\tilde P_2\to \tau \tilde P_3^{[k]}$ which is of dimension $(1,1)$. In view of Lemma \ref{lem:2equivalence}, we forgo to insert the corresponding rectangles around the subquiver and just connected the $2$-arrows to the sinks of the corresponding subquivers. \\

@@

Now we can recursively the $2$-quivers of $ \tilde P_m^{[k]}$ for $m\geq 4$ and $k=n-1,\ldots,0$ when repeating the procedure. For $m\geq 4$, the truncated preprojective $\tau\tilde P_m^{[k]}$ can uniquely be found as a quotient of $\tilde P_m$.
%Recall that we have $\udim P_m^{[k]}=\udim P_{m-1}^{[1]}+(n-k-1)\udim P_{m-1}$. 
As we have a one-to-one correspondence between basis elements of $\Hom(P_{m-1},\tau P_m^{[1]})$ and $\Ext(P_m^{[1]},P_{m-1})$ when passing to the universal covering quiver, we can proceed as before. Let us assume that we already constructed the $2$-quiver of $ P_{m+1}^{[k]}$. The $2$-quiver of $P^{[k-1]}_{m+1}$ is given by glueing the $2$-quiver of $ P_{m+1}^{[k]}$ to the one of $ P_{m}$
\\

@@

\noindent where the $2$-arrows connects the $2$-quiver of $\tilde P_{m+1}^{[k]}$ to its Auslander-Reiten translate $\tau \tilde P_{m+1}^{[k]}= \tilde P_{m-1}^{[k]}$ in $\tilde P_m$. This subquiver can be found at the very right of the $2$-quiver of $\tilde P_m$. Note that forgo the colouring of the arrows as it does not play any role from the combinatorial point of view. As an example consider the case $n=3$ and $m=3$. The $2$-quiver of $\tilde P_3$ is given by:\\

@@

Write $\mathcal Q(m,k)$ for the $2$-quiver of $\tilde P_m^{[k]}$. Clearly, every vertex  $q\in\mathcal Q(m,k)_0$ corresponds a vertex $\tilde q\in \widetilde{K(n)_0}$. Note that two different vertices of $\mathcal Q(m,k)_0$ can correspond to the same vertex of $\widetilde{K(n)}$. Writing dimension vectors $\tbfe\in\NN^{\widetilde{K(n)_0}}$ as $\tbfe=\sum_{q\in \widetilde{K(n)_0}}\tbfe_q\cdot q$, the dimension types $\tbfe (\beta)$ and  $\bfe (\beta)$ of a subset $\beta\subset\mathcal Q(m,k)_0$ are defined by
$$\tbfe (\beta)=\sum_{q\in\beta}\tilde q\in\NN^{\widetilde{K(n)_0}},\quad \bfe (\beta)=F_{K(n)}\tbfe (\beta).$$

\begin{theorem}\begin{enumerate}
\item  For $\tbfe\in\NN^{\widetilde{K(n)_0}}$, the Euler characteristic $\chi(\Gr_{\tbfe }(\tilde P_m^{[k]}))$ is given by the number of strong successor closed subsets of dimension type $\tbfe $ of the $2$-quiver of $\mathcal Q(m,k)$.
	\item For $\bfe\in\NN^{K(n)_0}$, the Euler characteristic $\chi(\Gr_{\bfe }(P_m^{[k]}))$ is given by the number of strong successor closed subsets of dimension type $\bfe $ of the $2$-quiver of $\mathcal Q(m,k)$.
\end{enumerate}
\end{theorem}
\begin{proof}
The results of Section \ref{QG} imply that the second statement follows from the first one.

We proceed by induction on $m$ and $k$. The strong successor closed subsets of the $2$-quivers of $\tilde P_1$, $\tilde P_2$ and $\tilde P_2^{[n-1]}$ coincide with the successor closed subsets of the quivers (\ref{coeff}) which we denote by $\Gamma_M$ for $M\in\{P_1, P_2,P_2^{[n-1]}\}$.
  Passing to the universal cover, we have $\Gr_e(\tilde M)\in\{\varnothing,\{\pt\}\}$ such that $\Gr_e(\tilde M)=\{\pt\}$ if and only if $e\subset (\Gamma_M)_0$ is successor closed.
  As the number cells of $\Gr_e(\tilde M)$ and $\Gr_e(M)$ coincide, the claim follows.

  Thus assume that the claim is true for $P_m$ (resp. $\tilde P_{m}$) and $P_{m+1}^{[k]}$ (resp. $\tilde P_{m+1}^{[k]}$) and consider the short exact sequence
  \[
    \ses{\tilde P_m}{\tilde P_{m+1}^{[k-1]}}{\tilde P_{m+1}^{[k]}}.
  \]
  Denote the corresponding $2$-quivers by $\Gamma_{P_m}$ and $\Gamma_{P_{m+1}^{[k]}}$ and let $\beta_1\cup\beta_2\subset (\Gamma_{P_m})_0\cup (\Gamma_{P_{m+1}^{[k]}})_0$ be a pair of strong successor closed subsets which gives rise to a pair of non-empty cell by induction hypothesis.
  Denote by $d(\beta_1+\beta_2)$ the dimension vector corresponding to this subset.
  By Proposition \ref{fibers} the pair $\beta_1\cup\beta_2$, it does not correspond to a non-empty cell of $\Gr_{d(\beta_1+\beta_2)}(\tilde P_{m+1}^{[k]})$ if and only if $\beta_2=(\Gamma_{P_{m+1}^{[k]}})_0$ and $\beta_1\cap(\Gamma_{\tau\tilde P_{m+1}^{[k]}})_0=\varnothing$.
  But this is precisely the condition on $\beta_1\cup\beta_2$ to be strong successor closed as $\Gamma_{P_{m+1}^{[k]}}$ is connected to $\Gamma_{\tau\tilde P_{m+1}^{[k]}}$ by a $2$-arrow.
\end{proof}

>>>>

%\begin{remark}\footnote{Insert a remark that we do not work on the universal cover} 
%  Lemma \ref{AR} can be generalized to the cases $m\leq 2$.
%  The only technical problem is that $\tau \tilde P^{(k)}_{m+1}$ is no quotient of $\tilde P_{m,k}$ and we have to replace it by the image of the unique morphism to $\tau \tilde P^{(k)}_{m+1}$ which induces the exact sequence. 
%\end{remark}
In the following, we frequently use the notation and conventions of Section \ref{Lifting}.
The next step is to associate a $2$-quiver $\mathcal Q(m,k)$ with the lifted truncated preprojective representations $\tilde P_m^{[k]}:=\tilde P_m^{\{1,\ldots,k\}}$ for $m\geq 2$, $0\leq k\leq n-1$  whose strong successor closed subsets are in one-to-one correspondence with cells of quiver Grassmannians. Recall that the cells of the quiver Grassmannians attached to $P_m^{[k]}:= F_{K(n)}\tilde P_m^{[k]}$ are in one-to-one correspondence with those attached to the lift $\tilde P_m^{[k]}$. Moreover, the quiver Grassmannians $\Gr_{\tbfe}(P_m^V)$ of truncated preprojective representations only depend on $\tbfe$ and $\udim P_m^V$. In particular, in view of Theorem \ref{thm:truncpp}, the $2$-quivers $\mathcal Q(m,k)$ can be seen as being attached to any (truncated) preprojective representation of this dimension.

@@

\begin{align}\label{eq:homext} \Hom(P_{m-1},\tau P_m^{[k]})\cong\bigoplus_{i=k+1}^n\Hom(\alpha_{i}^{(-1)^{m}}\tilde P_{m-1},\tau\tilde P_{m}^{[k]})\cong\bigoplus_{i=k+1}^n\Ext(\tilde P_m^{[k]},\alpha_{i}^{(-1)^{m}}\tilde P_{m-1})\cong\Ext(P_m^{[k]},P_{m-1}).\end{align}

@@

The representation $\tau\tilde P_3^{[k]}$ is given by 
\[\xymatrix@R10pt@C20pt{&&(2,\alpha_{k+1}^{-1})\ar_{\alpha_{k+1}}[lld]\\ (1,1) &&\vdots\\&&(2,\alpha_n^{-1})\ar_{\alpha_{n}}[llu]}.\]
Note that the image of the homomorphism of the corresponding exact sequence induced by the respective basis elements is the subquiver $(2,\alpha_i^{-1})\xrightarrow{\alpha_i} (1,1)$ for $i=k+1,\ldots, n$. Thus in this case the $2$-arrow connects the $2$-quiver of $\tilde P_3^{[k]}$ to the image of the corresponding homomorphism $\tilde P_2\to \tau \tilde P_3^{[k]}$ which is of dimension $(1,1)$. In view of Lemma \ref{lem:2equivalence}, we forgo to insert the corresponding rectangles around the subquiver and just connected the $2$-arrows to the sinks of the corresponding subquivers.

 The chosen basis of $\Ext(P_3^{[n-1]},P_2)$ recursively gives rise to the following $2$-quiver of $P_3^{[k]}$ for $k=n-1,\ldots,0$ where we drop the indices of the vertices in abuse of notation. Note that the middle vertex of $\tilde P_3^{[n-1]}$ is $(2,\alpha_n^{-1})$ and that all $2$-arrows end in the vertex corresponding to $(1,1)$ of $\widetilde{K(n)}$.
\\

@@

Now we can recursively the $2$-quivers of $ \tilde P_m^{[k]}$ for $m\geq 4$ and $k=n-1,\ldots,0$ when repeating the procedure and considering the natural basis of $\Ext(P_m^{[n-1]},P_{m-1})$ induced by the decomposition (\ref{eq:homext}). As we have a one-to-one correspondence between basis elements of $\Hom(P_{m-1},\tau P_m^{[k]})$ and $\Ext(P_m^{[k]},P_{m-1})$ when passing to the universal covering quiver, we can proceed as before.

For $m\geq 4$, the truncated preprojective $\tau\tilde P_m^{[k]}$ can uniquely be found as a quotient of $\tilde P_m$.
%Recall that we have $\udim P_m^{[k]}=\udim P_{m-1}^{[1]}+(n-k-1)\udim P_{m-1}$. 
 Let us assume that we already constructed the $2$-quiver of $ P_{m+1}^{[k]}$. The $2$-quiver of $\tilde P^{[k-1]}_{m+1}$ is given by glueing the $2$-quiver of $\tilde P_{m+1}^{[k]}$ to the one of $\tilde P_{m}$ where the $2$-arrows connects the $2$-quiver of $\tilde P_{m+1}^{[k]}$ to its Auslander-Reiten translate $\tau \tilde P_{m+1}^{[k]}= \tilde P_{m-1}^{[k]}$ in $\tilde P_m$. This subquiver can be found at the very right of the $2$-quiver of $\tilde P_m$.
\\

@@

\noindent In order to keep the diagrams as simple as possible we did not insert the labels of the vertices and arrows. As  every vertex  $q\in\mathcal Q(m,k)_0$ corresponds a vertex $\tilde q\in \widetilde{K(n)_0}$, in any case it is straightforward to associate an arrow $\alpha_i\in K(n)_1$ with every ($2$-)arrow of $\mathcal Q(m,k)$. As an example consider the case $n=3$ and $m=3$. The $2$-quiver of $\tilde P_3$ is given by:\\

@@

 Note that two different vertices of $\mathcal Q(m,k)_0$ can correspond to the same vertex of $\widetilde{K(n)}$. Writing dimension vectors $\tbfe\in\NN^{\widetilde{K(n)_0}}$ as $\tbfe=\sum_{q\in \widetilde{K(n)_0}}\tbfe_q\cdot q$, the dimension types $\tbfe (\beta)$ and  $\bfe (\beta)$ of a subset $\beta\subset\mathcal Q(m,k)_0$ are defined by
$$\tbfe (\beta)=\sum_{q\in\beta}\tilde q\in\NN^{\widetilde{K(n)_0}},\quad \bfe (\beta)=F_{K(n)}\tbfe (\beta).$$

\begin{theorem}\begin{enumerate}
\item  For $\tbfe\in\NN^{\widetilde{K(n)_0}}$, the Euler characteristic $\chi(\Gr_{\tbfe }(\tilde P_m^{[k]}))$ is given by the number of strong successor closed subsets of dimension type $\tbfe $ of the $2$-quiver of $\mathcal Q(m,k)$.
	\item For $\bfe\in\NN^{K(n)_0}$, the Euler characteristic $\chi(\Gr_{\bfe }(P_m^{[k]}))$ is given by the number of strong successor closed subsets of dimension type $\bfe $ of the $2$-quiver of $\mathcal Q(m,k)$.
\end{enumerate}
\end{theorem}
\begin{proof}
The results of Section \ref{QG} imply that the second statement follows from the first one.
We proceed by induction on $m$ and $k$. The case of the representation $\tilde P_1$ is trivial. The strong successor closed subsets of the $2$-quivers  $\mathcal Q(2,k)$ of $\tilde P_2^{[k]}$ coincide with the successor closed subsets of the quivers (\ref{coeff}). As $(\tilde P_2^{[k]})_q\in\{0,1\}$ for all $q\in\widetilde{K(n)}$, the  subrepresentations are in one-to-one correspondence with the successor closed subsets of (\ref{coeff}). Equivalently, we have $\Gr_{\tbfe}(\tilde P_2^{[k]})\in\{\varnothing,\{\pt\}\}$ such that $\Gr_{\tbfe}(\tilde P_2^{[k]})=\{\pt\}$ if and only if $e\subset \mathcal Q(2,k)$ is successor closed.

  Thus assume that the claim is true for $\tilde P_m$  and $\tilde P_{m+1}^{[k]}$ and consider the short exact sequence
  \[
    \ses{\tilde P_m}{\tilde P_{m+1}^{[k-1]}}{\tilde P_{m+1}^{[k]}}.
  \]
  We have $\mathcal Q(m+1,k-1)_0=\mathcal Q(m+1,k)_0\coprod{}\mathcal Q(m,0)_0$.
	Let $\beta_1\coprod\beta_2\subset \mathcal Q(m+1,k)_0\coprod{}\mathcal Q(m,0)_0$ be a pair of strong successor closed subsets which gives rise to a pair of non-empty cells by induction hypothesis.
 % Denote by $d(\beta_1+\beta_2)$ the dimension vector corresponding to this subset.
  By Proposition \ref{quotient}, the fibre over the pair of cells corresponding to $\beta_1\coprod\beta_2$ is empty if and only if $\beta_2=\mathcal Q(m+1,k)_0$ and $\beta_1\cap\mathcal T(m-1,k)_0=\varnothing$ where $\mathcal T(m-1,k)$ is the $2$-quiver of $\tilde\tau P_{m+1}^{[k]}=\tilde P_{m-1}^{[k]}$ sitting inside $\mathcal Q(m,0)$.
  But this is precisely the condition on $\beta_1\coprod\beta_2$ to be strong successor closed as $\mathcal Q(m+1,k)$ is connected to $\mathcal T(m-1,k)$ by a $2$-arrow.\sayT{compatibility? Classify the cells which satisfy the condition that the quotient on the AR-translate is zero...}
\end{proof}
